Approximate Bayesian Computation in Large Scale Structure: constraining the galaxy-halo connection

### 2017-03-14 ccppabc/text/revision/clean_manuscript.tex
clean manuscript

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[Hahn \& Vakili et al.]{
\parbox{\textwidth}{
ChangHoon~Hahn,$^{1,\ast}$\thanks{E-mail: chh327@nyu.edu}
Mohammadjavad~Vakili,$^{1,\ast}$\thanks{E-mail: mjvakili@nyu.edu}
Kilian~Walsh,$^{1}$\\
Andrew~P.~Hearin,$^{2}$
David~W.~Hogg,$^{1,3,4,5}$
and Duncan Campbell,$^{6}$
}
  \vspace*{10pt} \\
% List of institutions
$*$ These authors have contributed equally to the paper\\ 
$^{1}$Center for Cosmology and Particle Physics,Department of Physics, New York University, 726 Broadway, New York, NY\\
$^{2}$Yale Center for Astronomy \& Astrophysics, Yale University, New Haven, CT\\
$^{3}$Flatiron institute, 160 Fifth Avenue, New York, NY\\
$^{4}$Center for Data Science, New York University, 60 Fifth Ave, New York, NY 10011\\
$^{5}$Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117 Heidelberg, Germany\\
$^{6}$Department of Astronomy, Yale University, New Haven, CT
}

@@

% Abstract of the paper
\begin{abstract}
{\bf \color{dred}
Standard approaches to Bayesian parameter inference in large scale 
structure assume a Gaussian functional form (chi-squared form) 
for the likelihood. This assumption, in detail, cannot %They are also limited to analyzing measurements of the two-point correlation function. 
be correct. Likelihood free inferences such as Approximate Bayesian Computation (ABC) 
relax these restrictions and make inference possible without making any 
assumptions on the likelihood.
} Instead ABC relies on a forward generative model of the data and a metric 
for measuring the distance between the model and data. In this work, we 
demonstrate that ABC is feasible for LSS parameter inference by using it 
to constrain parameters of the halo occupation distribution (HOD) model 
for populating dark matter halos with galaxies.

@@

%%%%%%%%%%%%%%%%%%%%%%% SECTION: %%%%%%%%%%%%%%%%%%%%%%%
% SECTION: Comparison to MCMC analysis  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Comparison to the Gaussian {\color{dred} Pseudo - Likelihood} MCMC Analysis} \label{sec:abcvsmcmc}
In order to assess the quality of the parameter inference described in the previous section, 
we compare the ABC-PMC results with the HOD parameter constraints 
{\bf \color{dred} from}
assuming a Gaussian likelihood 
function. The model used for the Gaussian likelihood analysis is different than the forward 
generative model adopted for the ABC-PMC algorithm, to be consistent with the standard approach.

In the ABC analysis, the model accounts for sample variance by randomly sampling a subvolume to be 
populated with galaxies. 
{\bf \color{dred} %%%
Instead, in the Gaussian pseudo-likelihood analysis, the covariance matrix is assumed to capture the
uncertainties from sample variance. Hence, in the model for the Gaussian pseudo-likelihood analysis, 
we populate halos of the {\em entire} $\mathtt{MultiDark}$ simulation rather than a subvolume.
We describe the Gaussian pseudo-likelihood analysis below.
}
%The Gaussian likelihood analysis assumes a covariance matrix that captures the uncertainties from the sample variance, and therefore in the model, we populate the halos of the entire $\mathtt{MultiDark}$ simulation rather than a subvolume.

{\bf \color{dred}
To write down the Gaussian pseudo-likelihood, we first introduce the vector $\mathbf{x}$: 
a combination of the summary statistics (observables) for a galaxy catalog. 
When we use $\ngalaxy$ and $\xigg(r)$ as observables in the analysis: $\mathbf{x} = [\ngalaxy , \xigg]$;
when we use $\ngalaxy$ and $\gmf(N)$ as observables in the analysis: $\mathbf{x} = [\ngalaxy , \gmf]$.
Based on this notation, we can write pseudo-likelihood function as 
\begin{eqnarray}
-2 \ln \mathcal{L}(\theta | d) &=& \Delta \mathbf{x}^{T}\widehat{C^{-1}}\Delta \mathbf{x} + \ln\Big[(2\pi)^{d}\mathrm{det}(C)\Big], \label{loglike}
\end{eqnarray}
where 
\begin{eqnarray}
\Delta \mathbf{x} &=& [\mathbf{x}_{obs} -\mathbf{x}_{mod}], 
\end{eqnarray}
the difference between $\mathbf{x}_{obs}$, measured from the mock observation, 
and $\mathbf{x}_{mod}(\mathbb{\theta})$ measured from the mock catalog generated 
from the model with parameters $\theta$ .
$d$ here is the dimension of $\mathbf{x}$ (for $\mathbf{x} = [\ngalaxy, \xigg]$, $d = 13$; 
for $\mathbf{x} = [\ngalaxy, \gmf]$, $d = 10$).  
$\widehat{C^{-1}}$ is the inverse covariance matrix, which we estimate following \cite{hartlap2007}:
\begin{eqnarray}
\widehat{C^{-1}} = \frac{N_{\rm mocks}-d - 1}{N_{\rm mocks} -1} \; \widehat{C}^{-1}.
\end{eqnarray}
$\widehat{C}$ is the estimated covariance matrix, calculated using the corresponding 
$\mathbf{x}$ block of the covariance matrix from Eq.~\ref{eq:cov}, and $N_{\rm mock}$ 
is the number of mocks used for the estimation ($N_{\rm mock} = 124$; see Section \ref{sec:mock_obv}).
We note that in $\widehat{C}$ the dependence on the HOD parameters is neglected, 
so the second term in the expression of Eq.~\ref{loglike} can be neglected. 
Finally, using this pseudo-likelihood, we sample from the posterior distribution 
given the prior distribution using the MCMC sampler $\mathtt{emcee}$ (\citealt{emcee}). 
}

{\bf \color{dred}
In Figures~\ref{fig:hist_nbarxi} and \ref{fig:hist_nbargmf}, we compare the results from
ABC-PMC and Gaussian pseudo-likelihood MCMC analyses using $[\ngalaxy, \xigg]$ and 
$[\ngalaxy, \gmf]$ as observables, respectively. The top panels in each figure 
compares the marginalized posterior PDFs for three parameters of the HOD model: 
$\{\log \mathcal{M}_{\rm min}, \alpha, \log\mathcal{M}_1\}$. The lower panels in each 
figure compares the $68\%$ and $95\%$ confidence intervals of the constraints 
derived from the two inference methods as a box plot. The ``true'' HOD parameters
are marked by vertical dashed lines in each panel. 
}

{\bf \color{dred}
In both Figures~\ref{fig:hist_nbarxi} and \ref{fig:hist_nbargmf}, the marginalized 
posteriors for each of the parameters from both inference methods are comparable 
and consistent with the ``true'' HOD parameters. However, we note that there are 
minor discrepancies between the maringalized posterior distributions. In particular, 
the distribution for $\alpha$ derived from ABC-PMC is less biased than the $\alpha$ 
constraints from the Gaussian pseudo-likelihood approach. 
}
%The uncertainty over the parameter $\log \mathcal{M}_\mathrm{min}$ is slightly smaller while the uncertainties of the parameters $\alpha$ and $\log M_{1}$ are slightly larger for our ABC-PMC results.  In Figure \ref{fig:hist_nbargmf}, when we use $\ngalaxy$ and $\gmf$ as observables, both methods produce consistent constraints with each other and ``true'' HOD parameters. The lower panel of Figure \ref{fig:hist_nbargmf} shows that the uncertainties from the two methods are comparable. The ABC-PMC constraints over $\alpha$ is slightly less biased and slightly less precise.

{\bf \color{dred}
In Figures \ref{fig:cont_nbarxi} and \ref{fig:cont_nbargmf}, we plot the contours 
enclosing the $68\%$ and $95\%$ confidence regions of the posterior probabilities of 
the two methods using $[\ngalaxy, \xigg]$ and $[\ngalaxy, \gmf]$ as observables 
respectively. In both figures, we mark the ``true'' HOD parameters (black star). The 
overall shape of the contours are in agreement with each other. However, we note
that the contours for the ABC-PMC method are more extended along $\alpha$. 
}

%%% RefReport #1 
%%% CHH (3/8/2017): Made significant changes to the text in order to better emphasize
%%%      the advantages of ABC over MCMC
{\bf \color{dred} 
Overall, the HOD parameter constraints from ABC-PMC are consistent with those from 
the Gaussian pseudo-likelihood MCMC method; however, using ABC-PMC has a number of 
advantages. For instance, ABC-PMC utilizes a forward generative model. Our forward 
generative model accounts for sample variance. On the other hand, the Gaussian 
pseudo-likelihood approach, as mentioned earlier this section, does not account 
for sample variance in the model and relies on the covariance matrix estimate to 
capture the sample variance of the data. 
}

{\bf \color{dred} 
Accurate estimation of the covariance matrix 
in LSS, however, faces a number of challenges. It is both labor and computationally 
expensive and dependent on the accuracy of simulated mock catalogs, known to be 
unreliable on small scales~(see \citealt{cosmiccode,nifty} and references therein). 
In fact, as \cite{Sellentin:2016a} points out, using estimates of the covariance 
matrix in the Gaussian psuedo-likelihood approach become further problematic. Even 
when inferring parameters from a Gaussian-distributed data set, using covariance 
matrix estimates rather than the {\em true} covariance matrix leads to a likelihood 
function that is {\em no longer} Gaussian. ABC-PMC does not depend on a covariance 
matrix estimate; hence, it does not face these problems.
}

{\bf \color{dred}
In addition to not requiring accurate covariance matrix estimates, forward models 
of the ABC-PMC method, in principle, also have the advantage that they can account 
for sources of systematic uncertainties that affect observations. All observations 
suffer from significant systematic effects which are often difficult to correct. 
For instance, in SDSS-III BOSS~\citep{Dawson:2013a}, fiber collisions and redshift
failures siginifcantly bias measurements and analysis of observables such as 
$\xigg$ or the galaxy powerspectrum~\citep{Ross:2012aa, Guo:2012a, Hahn:2017a}. 
In parameter inference, these systematics can affect the likelihood, and thus any 
analysis that requires writing down the likelihood, in unknown ways. With a forward generative model of the ABC-PMC 
method, the systematics can be simulated and marginalized out to achieve unbiased 
constraints. 
} 
%{\bf \color{darkgreen} CHH: @MJV I incorporated the sentence into the paragraph. Can you check it out and confirm it's fine?  } 
%In general, we find that the constraints from ABC-PMC are consistent with those from the Gaussian-likelihood MCMC method both in terms of accuracy and precision. However, we emphasize some advantages of using ABC-PMC. ABC-PMC utilizes a forward generative model. Our forward generative model, for instance, accounts for sample variance. In principle, they also have the advantage that they can account for sources of systematic uncertainties in the observational data. 
%Furthermore, the Gaussian-likelihood  approach relies on constructing an accurate covariance matrix estimate that captures the sample variance of the data. While we are able to do this accurately within the scope of the HOD framework, for more general LSS parameter inference situations, it is both labor and computationally expensive and dependent on the accuracy of simulated mock catalogs, which are known to be unreliable on small scales (see \citealt{cosmiccode,nifty} and references therein).  Since ABC-PMC utilizes a forward model to account for sample variance, it does not depend on a covariance matrix estimate; hence it does not face these problems.

{\bf \color{dred}
Furthermore, {\em ABC-PMC -- unlike the Gaussian pseudo-likelihood approach -- 
is agnostic about the functional form of the underlying distribution of the 
summary statistics} (\emph{e.g.} $\xigg$ and $\gmf$). As we explain throughout 
the paper, the likelihood function in LSS {\em cannot} be Gaussian. For $\xigg$, 
the correlation function must satisfy non-trivial positive-definiteness requirements 
and hence the Gaussian pseudo-likelihood function assumption is not correct 
in detail. In the case of $\gmf(N)$, assuming a Gaussian functional form for 
the likelihood, which in reality is more likely Poisson, misrepresents the true 
likelihood function. In fact, this incorrect 
likelihood, may explain why the constraints on $\alpha$ are less biased for 
the ABC-PMC analysis than the Gaussian-likelihood analysis in \ref{fig:cont_nbargmf}.
}
%{\bf \color{darkgreen} CHH: @MJV, I parrot a lot of arguments made in the intro without providing much extra 'meat'. Let me know what you think.  }

{\bf \color{dred} 
Although in our comparison using simple mock observations, we find generally 
consistent parameter constraints from both the ABC-PMC analysis and the standard 
Gaussian pseudo-likelihood analysis, more realistic scenarios present many factors
that can generate inconsistencies. Consider a typical galaxy catalog from 
LSS observations. These catalogs consist of objects with different data 
qualities, signal-to-noise ratios, and systematic effects. For example, catalogs 
are often incomplete beyond some luminosity/redshift or have some threshold 
signal-to-noise ratio cut imposed on them.
}

%It is possible to consider a scenario in which the results from the standard analysis are inconsistent with the results from the ABC.  Let us consider a dataset---from a galaxy survey---consisting of objects with different data qualities, signal-to-noise ratios, and systematic effects.  This could be a catalog of galaxies which is incomplete beyond a certain luminosity or redshift. Another example could be a catalog with a signal-to-noise ratio cut.
{\bf \color{dred} 
These selection effects, coupled with the systematic effects earlier this section, 
make correctly predicting the likelihood intractable. In the standard Gaussian 
pseudo-likelihood analysis, and other analysis that require writing down a 
likelihood function, these effects can significantly 
bias the inferred parameter constraints. In these situations, employing
ABC equipped with a generative forward model that incorproates selection and 
systematic effects may produce less biased parameter constraints. 
}
%Methods that rely on a likelihood function may not correctly predict how the functional form of the likelihood is impacted by the presence of possible selection effects and systematic uncertainties.  
%Alternatively, one could alleviate this issue by employing ABC equipped with a generative forward model of the simulated data that incorporates signal-to-noise ratio cuts, incompleteness, systematic uncertainties etc. }{
%\bf \color{darkgreen} MJV: @CHH Can you read the two paragraphs above and see if they address referee's concern about the extreme data situation that could lead to discrepancy ? }
%%%
%In general, we find that the constraints from ABC-PMC are consistent with those from the Gaussian-likelihood 
%MCMC method both in terms of accuracy and precision. However, we emphasize some advantages of using 
%ABC-PMC. ABC-PMC utilizes a forward generative model. Our forward generative model, for instance, 
%accounts for sample variance. In principle, they also have the advantage that they can account for 
%sources of systematic uncertainties in the observational data.

{\bf \color{dred}
Despite the advantages of ABC, one obstactle for adopting it to parameter 
inference has been the computational costs of generative forward models, a 
key element of ABC. By combining ABC with the PMC sampling method, however, 
ABC-PMC efficiently converges to give reliable posterior parameter constraints. 
In fact, in our analysis, the total computational resources required for the 
ABC-PMC analysis were {\em comparable} to the computational resources used 
for the Gaussian pseudo-likelihood analysis with MCMC sampling.
}

{\bf \color{dred}
Applying ABC-PMC beyond the analysis in this work, to broader LSS analyses 
imposes some caveats. In this work, we focus on the galaxy-halo connection, so
our generative forward model populates halos with galaxies. LSS analyses 
for inferring cosmological parameters would require generating halos by 
running cosmological simulations. The forward models also need to accurately 
model the observation systematic effects of the latest observations. Hence, 
accurate generative forward models in LSS analyses demand improvements in simulations 
and significant computational resources in order to infer unbiased parameter 
constraints. Recent cosmology simulations show promising improvements 
in both accuracy and speed~\citep[\emph{e.g.}][]{fastpm}. Such developements
will be crucial for applying ABC-PMC to broader LSS analyses and exploiting 
the significant advantages that ABC-PMC offers.
}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ABC vs MCMC Histograms 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
%\begin{center}
\includegraphics[width=1.\textwidth]{figs/paper_ABCvsMCMC_nbarxi.pdf}
\caption{\label{fig:hist_nbarxi} 
{\bf \color{dred}
We compare the $\log \mathcal{M}_{\rm min}$, $\alpha$, and $\log \mathcal{M}_{1}$ parameter 
constraints from ABC-PMC (orange) to constraints from the Gaussian pseudo-ikelihood MCMC (blue) 
using $\ngalaxy$ and $\xigg(r)$ as observables. The \emph{top} panels compares the two methods' 
marginalized posterior PDFs over the parameters. In the \emph{bottom} panels, we include box 
plots marking the confidence intervals of the posterior distributions. The boxes represent the 
$68\%$ confidence interval while the ``whiskers'' represent the $95\%$ confidence interval. We mark 
the ``true'' HOD parameters with vertical black dashed line. The marginalized posterior PDFs 
obtained from the two methods are consistent with each other. The ABC-PMC and Gaussian pseudo-likelihood
constraints are generally consistent for $\log \mathcal{M}_{\rm min}$ and $\log \mathcal{M}_{1}$. 
The ABC-PMC constraint for $\alpha$ is slightly less biased and has slightly larger uncertainty
then the constraint from Gaussian pseudo-likelihood analysis.} 
}
%\end{center}
\end{figure*}

\begin{figure*}
%\begin{center}
\includegraphics[width=1.\textwidth]{figs/paper_ABCvsMCMC_nbargmf.pdf}
\caption{\label{fig:hist_nbargmf} 
{\bf \color{dred}
Same as Figure \ref{fig:hist_nbarxi}, but both the ABC-PMC
analysis and the Gaussian pseudo-likelihood MCMC analysis use $\ngalaxy$ and $\gmf(N)$ as 
observables. Both methods derive constraints consistent with the ``true'' HOD parameters 
and infer the region of allowed values to similar precision. We note that the MCMC constraint 
on $\alpha$ is slightly more biased compared to ABC-PMC estimate. This discrepancy may stem 
from the fact that the use of Gaussian pseudo-likelihood and its associated assumptions is more 
spurious when modeling the group multiplicity function.}
}
%\end{center}
\end{figure*}

>>>>

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[Hahn \& Vakili et al.]{
\parbox{\textwidth}{
ChangHoon~Hahn,$^{1,\ast}$\thanks{E-mail: chh327@nyu.edu}
Mohammadjavad~Vakili,$^{1,\ast}$\thanks{E-mail: mjvakili@nyu.edu}
Kilian~Walsh,$^{1}$\\
Andrew~P.~Hearin,$^{2}$
David~W.~Hogg,$^{1,3,4,5}$
and Duncan Campbell,$^{6}$
}
  \vspace*{10pt} \\
% List of institutions
$*$ These authors have contributed equally to the paper\\ 
$^{1}$Center for Cosmology and Particle Physics, Department of Physics, New York University, 726 Broadway, New York, NY\\
$^{2}$Yale Center for Astronomy \& Astrophysics, Yale University, New Haven, CT\\
$^{3}$Flatiron institute, 160 Fifth Avenue, New York, NY\\
$^{4}$Center for Data Science, New York University, 60 Fifth Ave, New York, NY 10011\\
$^{5}$Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117 Heidelberg, Germany\\
$^{6}$Department of Astronomy, Yale University, New Haven, CT
}

@@

% Abstract of the paper
\begin{abstract}

Standard approaches to Bayesian parameter inference in large scale 
structure assume a Gaussian functional form (chi-squared form) 
for the likelihood. This assumption, in detail, cannot %They are also limited to analyzing measurements of the two-point correlation function. 
be correct. Likelihood free inferences such as Approximate Bayesian Computation (ABC) 
relax these restrictions and make inference possible without making any 
assumptions on the likelihood. Instead ABC relies on a forward generative model of the data and a metric 
for measuring the distance between the model and data. In this work, we 
demonstrate that ABC is feasible for LSS parameter inference by using it 
to constrain parameters of the halo occupation distribution (HOD) model 
for populating dark matter halos with galaxies.

@@

%%%%%%%%%%%%%%%%%%%%%%% SECTION: %%%%%%%%%%%%%%%%%%%%%%%
% SECTION: Comparison to MCMC analysis  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Comparison to the Gaussian Pseudo-Likelihood MCMC Analysis} \label{sec:abcvsmcmc}
In order to assess the quality of the parameter inference described in the previous section, 
we compare the ABC-PMC results with the HOD parameter constraints from
assuming a Gaussian likelihood 
function. The model used for the Gaussian likelihood analysis is different than the forward 
generative model adopted for the ABC-PMC algorithm, to be consistent with the standard approach.

In the ABC analysis, the model accounts for sample variance by randomly sampling a subvolume to be 
populated with galaxies. Instead, in the Gaussian pseudo-likelihood analysis, the covariance matrix is assumed to capture the
uncertainties from sample variance. Hence, in the model for the Gaussian pseudo-likelihood analysis, 
we populate halos of the {\em entire} $\mathtt{MultiDark}$ simulation rather than a subvolume.
We describe the Gaussian pseudo-likelihood analysis below.

%The Gaussian likelihood analysis assumes a covariance matrix that captures the uncertainties from the sample variance, and therefore in the model, we populate the halos of the entire $\mathtt{MultiDark}$ simulation rather than a subvolume.

To write down the Gaussian pseudo-likelihood, we first introduce the vector $\mathbf{x}$: 
a combination of the summary statistics (observables) for a galaxy catalog. 
When we use $\ngalaxy$ and $\xigg(r)$ as observables in the analysis: $\mathbf{x} = [\ngalaxy , \xigg]$;
when we use $\ngalaxy$ and $\gmf(N)$ as observables in the analysis: $\mathbf{x} = [\ngalaxy , \gmf]$.
Based on this notation, we can write pseudo-likelihood function as 
\begin{eqnarray}
-2 \ln \mathcal{L}(\theta | d) &=& \Delta \mathbf{x}^{T}\widehat{C^{-1}}\Delta \mathbf{x} + \ln\Big[(2\pi)^{d}\mathrm{det}(C)\Big], \label{loglike}
\end{eqnarray}
where 
\begin{eqnarray}
\Delta \mathbf{x} &=& [\mathbf{x}_{obs} -\mathbf{x}_{mod}], 
\end{eqnarray}
the difference between $\mathbf{x}_{obs}$, measured from the mock observation, 
and $\mathbf{x}_{mod}(\mathbb{\theta})$ measured from the mock catalog generated 
from the model with parameters $\theta$ .
$d$ here is the dimension of $\mathbf{x}$ (for $\mathbf{x} = [\ngalaxy, \xigg]$, $d = 13$; 
for $\mathbf{x} = [\ngalaxy, \gmf]$, $d = 10$).  
$\widehat{C^{-1}}$ is the inverse covariance matrix, which we estimate following \cite{hartlap2007}:
\begin{eqnarray}
\widehat{C^{-1}} = \frac{N_{\rm mocks}-d - 1}{N_{\rm mocks} -1} \; \widehat{C}^{-1}.
\end{eqnarray}
$\widehat{C}$ is the estimated covariance matrix, calculated using the corresponding 
$\mathbf{x}$ block of the covariance matrix from Eq.~\ref{eq:cov}, and $N_{\rm mock}$ 
is the number of mocks used for the estimation ($N_{\rm mock} = 124$; see Section \ref{sec:mock_obv}).
We note that in $\widehat{C}$ the dependence on the HOD parameters is neglected, 
so the second term in the expression of Eq.~\ref{loglike} can be neglected. 
Finally, using this pseudo-likelihood, we sample from the posterior distribution 
given the prior distribution using the MCMC sampler $\mathtt{emcee}$ (\citealt{emcee}).

In Figures~\ref{fig:hist_nbarxi} and \ref{fig:hist_nbargmf}, we compare the results from
ABC-PMC and Gaussian pseudo-likelihood MCMC analyses using $[\ngalaxy, \xigg]$ and 
$[\ngalaxy, \gmf]$ as observables, respectively. The top panels in each figure 
compares the marginalized posterior PDFs for three parameters of the HOD model: 
$\{\log \mathcal{M}_{\rm min}, \alpha, \log\mathcal{M}_1\}$. The lower panels in each 
figure compares the $68\%$ and $95\%$ confidence intervals of the constraints 
derived from the two inference methods as a box plot. The ``true'' HOD parameters
are marked by vertical dashed lines in each panel.

In both Figures~\ref{fig:hist_nbarxi} and \ref{fig:hist_nbargmf}, the marginalized 
posteriors for each of the parameters from both inference methods are comparable 
and consistent with the ``true'' HOD parameters. However, we note that there are 
minor discrepancies between the maringalized posterior distributions. In particular, 
the distribution for $\alpha$ derived from ABC-PMC is less biased than the $\alpha$ 
constraints from the Gaussian pseudo-likelihood approach.

%The uncertainty over the parameter $\log \mathcal{M}_\mathrm{min}$ is slightly smaller while the uncertainties of the parameters $\alpha$ and $\log M_{1}$ are slightly larger for our ABC-PMC results.  In Figure \ref{fig:hist_nbargmf}, when we use $\ngalaxy$ and $\gmf$ as observables, both methods produce consistent constraints with each other and ``true'' HOD parameters. The lower panel of Figure \ref{fig:hist_nbargmf} shows that the uncertainties from the two methods are comparable. The ABC-PMC constraints over $\alpha$ is slightly less biased and slightly less precise.

In Figures \ref{fig:cont_nbarxi} and \ref{fig:cont_nbargmf}, we plot the contours 
enclosing the $68\%$ and $95\%$ confidence regions of the posterior probabilities of 
the two methods using $[\ngalaxy, \xigg]$ and $[\ngalaxy, \gmf]$ as observables 
respectively. In both figures, we mark the ``true'' HOD parameters (black star). The 
overall shape of the contours are in agreement with each other. However, we note
that the contours for the ABC-PMC method are more extended along $\alpha$.

%%% RefReport #1 
%%% CHH (3/8/2017): Made significant changes to the text in order to better emphasize
%%%      the advantages of ABC over MCMC

Overall, the HOD parameter constraints from ABC-PMC are consistent with those from 
the Gaussian pseudo-likelihood MCMC method; however, using ABC-PMC has a number of 
advantages. For instance, ABC-PMC utilizes a forward generative model. Our forward 
generative model accounts for sample variance. On the other hand, the Gaussian 
pseudo-likelihood approach, as mentioned earlier this section, does not account 
for sample variance in the model and relies on the covariance matrix estimate to 
capture the sample variance of the data.

Accurate estimation of the covariance matrix 
in LSS, however, faces a number of challenges. It is both labor and computationally 
expensive and dependent on the accuracy of simulated mock catalogs, known to be 
unreliable on small scales~(see \citealt{cosmiccode,nifty} and references therein). 
In fact, as \cite{Sellentin:2016a} points out, using estimates of the covariance 
matrix in the Gaussian psuedo-likelihood approach become further problematic. Even 
when inferring parameters from a Gaussian-distributed data set, using covariance 
matrix estimates rather than the {\em true} covariance matrix leads to a likelihood 
function that is {\em no longer} Gaussian. ABC-PMC does not depend on a covariance 
matrix estimate; hence, it does not face these problems.

In addition to not requiring accurate covariance matrix estimates, forward models 
of the ABC-PMC method, in principle, also have the advantage that they can account 
for sources of systematic uncertainties that affect observations. All observations 
suffer from significant systematic effects which are often difficult to correct. 
For instance, in SDSS-III BOSS~\citep{Dawson:2013a}, fiber collisions and redshift
failures siginifcantly bias measurements and analysis of observables such as 
$\xigg$ or the galaxy powerspectrum~\citep{Ross:2012aa, Guo:2012a, Hahn:2017a}. 
In parameter inference, these systematics can affect the likelihood, and thus any 
analysis that requires writing down the likelihood, in unknown ways. With a forward generative model of the ABC-PMC 
method, the systematics can be simulated and marginalized out to achieve unbiased 
constraints.

%{\bf \color{darkgreen} CHH: @MJV I incorporated the sentence into the paragraph. Can you check it out and confirm it's fine?  } 
%In general, we find that the constraints from ABC-PMC are consistent with those from the Gaussian-likelihood MCMC method both in terms of accuracy and precision. However, we emphasize some advantages of using ABC-PMC. ABC-PMC utilizes a forward generative model. Our forward generative model, for instance, accounts for sample variance. In principle, they also have the advantage that they can account for sources of systematic uncertainties in the observational data. 
%Furthermore, the Gaussian-likelihood  approach relies on constructing an accurate covariance matrix estimate that captures the sample variance of the data. While we are able to do this accurately within the scope of the HOD framework, for more general LSS parameter inference situations, it is both labor and computationally expensive and dependent on the accuracy of simulated mock catalogs, which are known to be unreliable on small scales (see \citealt{cosmiccode,nifty} and references therein).  Since ABC-PMC utilizes a forward model to account for sample variance, it does not depend on a covariance matrix estimate; hence it does not face these problems.

Furthermore, {\em ABC-PMC -- unlike the Gaussian pseudo-likelihood approach -- 
is agnostic about the functional form of the underlying distribution of the 
summary statistics} (\emph{e.g.} $\xigg$ and $\gmf$). As we explain throughout 
the paper, the likelihood function in LSS {\em cannot} be Gaussian. For $\xigg$, 
the correlation function must satisfy non-trivial positive-definiteness requirements 
and hence the Gaussian pseudo-likelihood function assumption is not correct 
in detail. In the case of $\gmf(N)$, assuming a Gaussian functional form for 
the likelihood, which in reality is more likely Poisson, misrepresents the true 
likelihood function. In fact, this incorrect 
likelihood, may explain why the constraints on $\alpha$ are less biased for 
the ABC-PMC analysis than the Gaussian-likelihood analysis in \ref{fig:cont_nbargmf}.

%{\bf \color{darkgreen} CHH: @MJV, I parrot a lot of arguments made in the intro without providing much extra 'meat'. Let me know what you think.  }

Although in our comparison using simple mock observations, we find generally 
consistent parameter constraints from both the ABC-PMC analysis and the standard 
Gaussian pseudo-likelihood analysis, more realistic scenarios present many factors
that can generate inconsistencies. Consider a typical galaxy catalog from 
LSS observations. These catalogs consist of objects with different data 
qualities, signal-to-noise ratios, and systematic effects. For example, catalogs 
are often incomplete beyond some luminosity/redshift or have some threshold 
signal-to-noise ratio cut imposed on them.

%It is possible to consider a scenario in which the results from the standard analysis are inconsistent with the results from the ABC.  Let us consider a dataset---from a galaxy survey---consisting of objects with different data qualities, signal-to-noise ratios, and systematic effects.  This could be a catalog of galaxies which is incomplete beyond a certain luminosity or redshift. Another example could be a catalog with a signal-to-noise ratio cut.

These selection effects, coupled with the systematic effects earlier this section, 
make correctly predicting the likelihood intractable. In the standard Gaussian 
pseudo-likelihood analysis, and other analysis that require writing down a 
likelihood function, these effects can significantly 
bias the inferred parameter constraints. In these situations, employing
ABC equipped with a generative forward model that incorproates selection and 
systematic effects may produce less biased parameter constraints.

%Methods that rely on a likelihood function may not correctly predict how the functional form of the likelihood is impacted by the presence of possible selection effects and systematic uncertainties.  
%Alternatively, one could alleviate this issue by employing ABC equipped with a generative forward model of the simulated data that incorporates signal-to-noise ratio cuts, incompleteness, systematic uncertainties etc. }{
%\bf \color{darkgreen} MJV: @CHH Can you read the two paragraphs above and see if they address referee's concern about the extreme data situation that could lead to discrepancy ? }
%%%
%In general, we find that the constraints from ABC-PMC are consistent with those from the Gaussian-likelihood 
%MCMC method both in terms of accuracy and precision. However, we emphasize some advantages of using 
%ABC-PMC. ABC-PMC utilizes a forward generative model. Our forward generative model, for instance, 
%accounts for sample variance. In principle, they also have the advantage that they can account for 
%sources of systematic uncertainties in the observational data.

Despite the advantages of ABC, one obstactle for adopting it to parameter 
inference has been the computational costs of generative forward models, a 
key element of ABC. By combining ABC with the PMC sampling method, however, 
ABC-PMC efficiently converges to give reliable posterior parameter constraints. 
In fact, in our analysis, the total computational resources required for the 
ABC-PMC analysis were {\em comparable} to the computational resources used 
for the Gaussian pseudo-likelihood analysis with MCMC sampling.

Applying ABC-PMC beyond the analysis in this work, to broader LSS analyses 
imposes some caveats. In this work, we focus on the galaxy-halo connection, so
our generative forward model populates halos with galaxies. LSS analyses 
for inferring cosmological parameters would require generating halos by 
running cosmological simulations. The forward models also need to accurately 
model the observation systematic effects of the latest observations. Hence, 
accurate generative forward models in LSS analyses demand improvements in simulations 
and significant computational resources in order to infer unbiased parameter 
constraints. Recent cosmology simulations show promising improvements 
in both accuracy and speed~\citep[\emph{e.g.}][]{fastpm}. Such developements
will be crucial for applying ABC-PMC to broader LSS analyses and exploiting 
the significant advantages that ABC-PMC offers.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ABC vs MCMC Histograms 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
%\begin{center}
\includegraphics[width=1.\textwidth]{figs/paper_ABCvsMCMC_nbarxi.pdf}
\caption{\label{fig:hist_nbarxi} 
We compare the $\log \mathcal{M}_{\rm min}$, $\alpha$, and $\log \mathcal{M}_{1}$ parameter 
constraints from ABC-PMC (orange) to constraints from the Gaussian pseudo-ikelihood MCMC (blue) 
using $\ngalaxy$ and $\xigg(r)$ as observables. The \emph{top} panels compares the two methods' 
marginalized posterior PDFs over the parameters. In the \emph{bottom} panels, we include box 
plots marking the confidence intervals of the posterior distributions. The boxes represent the 
$68\%$ confidence interval while the ``whiskers'' represent the $95\%$ confidence interval. We mark 
the ``true'' HOD parameters with vertical black dashed line. The marginalized posterior PDFs 
obtained from the two methods are consistent with each other. The ABC-PMC and Gaussian pseudo-likelihood
constraints are generally consistent for $\log \mathcal{M}_{\rm min}$ and $\log \mathcal{M}_{1}$. 
The ABC-PMC constraint for $\alpha$ is slightly less biased and has slightly larger uncertainty
then the constraint from Gaussian pseudo-likelihood analysis.} 
%\end{center}
\end{figure*}

\begin{figure*}
%\begin{center}
\includegraphics[width=1.\textwidth]{figs/paper_ABCvsMCMC_nbargmf.pdf}
\caption{\label{fig:hist_nbargmf} 
Same as Figure \ref{fig:hist_nbarxi}, but both the ABC-PMC
analysis and the Gaussian pseudo-likelihood MCMC analysis use $\ngalaxy$ and $\gmf(N)$ as 
observables. Both methods derive constraints consistent with the ``true'' HOD parameters 
and infer the region of allowed values to similar precision. We note that the MCMC constraint 
on $\alpha$ is slightly more biased compared to ABC-PMC estimate. This discrepancy may stem 
from the fact that the use of Gaussian pseudo-likelihood and its associated assumptions is more 
spurious when modeling the group multiplicity function.}
%\end{center}
\end{figure*}
